---
title: "music recommendation system"
author: "LI Han"
date: "6/3/2020"
output: html_document
---

### Part 1. EDA

```{r Loading Packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(feather)
library(data.table)
library(viridis)
library(DT)
library(lubridate)
library(magrittr)
options(tibble.print_max = 5, tibble.print_min = 5)
```


  
#### TRAIN set


```{r train read, message=FALSE, warning=FALSE, results=FALSE, echo=FALSE}
setwd('/Volumes/EricSeagate/HKUST/Term 2/Machine Learning/Individual Project')
train <- as.tibble(fread('train.csv'))
```



```{r show train, message=FALSE,result='asis', warning=FALSE, echo=FALSE}
train
#datatable(head(train,5), style="bootstrap", class="table-condensed", options = list(dom = 'tp',scrollX = TRUE))
```

Defining useful functions , 

```{r train - useful functions}
## ggplot setting for readable labels
readable_labs <- theme(axis.text=element_text(size=12),
                     axis.title=element_text(size=14),
                     plot.title = element_text(hjust = 0.5))

# Function to dislpay count of each category of the column and plot how it affects target
target_vs_column <-function(df, col_name, x , y, title)
                  {
  
                  temp_df <- df %>% 
                          group_by_(col_name) %>% 
                          summarize(count = n(), mean_target = mean(target)) %>% 
                          arrange(desc(mean_target)) 
                  
                  df_plot <- temp_df %>%  
                            ggplot(aes_string(col_name, "mean_target")) + 
                            geom_col(aes(fill=count)) +
                            scale_fill_gradient(low='turquoise', high = 'violet')+
                            coord_flip() +
                            labs(x = x,
                                 y = y,
                                 title= title) +
                            readable_labs
                          
                  print(df_plot)
                  return (temp_df)
                  
                  }

# Function to group songs and user by count and check it agains mean_target
target_vs_colcount <- function(df, col_name, x, y, title)
                    { 
  
                    df %>% 
                      group_by_(col_name) %>% 
                      summarize(count = n(), mean_target = mean(target)) %>% 
                      group_by(count) %>% 
                      summarize(new_count = n(), avg_target = mean(mean_target)) %>% 
                      rename(no_of_items = new_count, occurence = count) %>% 
                      print %>% 
                      arrange(desc(avg_target)) %>% 
                      print %>% 
                      ggplot(aes(occurence, avg_target)) +
                        geom_line(color='turquoise') +
                        geom_smooth(color='turquoise') +
                        labs(x = x,
                             y = y,
                             title= title) +
                        readable_labs

                  
                  }

```

##### Train column count and its effect on target {.tabset}

##### source_system_tab


```{r sst}
target_vs_column(train, col_name = "source_system_tab",
                  x = 'Frequency',
                  y = 'Target',
                  title = 'Count of source_system_tab vs Target')
```

##### source_screen_name



```{r ssn}
target_vs_column(train, col_name = "source_screen_name",
                  x = 'Frequency',
                  y = 'Target',
                  title = 'Count of source_screen_name vs Target')
```

##### source_type
 
```{r st}
target_vs_column(train, col_name = "source_type",
                  x = 'Frequency',
                  y = 'Target',
                  title = 'Count of source_type vs Target')
```

#### Song count and User count vs target {.tabset}


##### Song count vs Target


```{r song count, message=FALSE, warning=FALSE}
target_vs_colcount(train, "song_id", "Song Occurence", "Target", "Song Occurence vs Target")
```

##### User count vs Target

If the user occurs more in the train data set (frequent listener) then it does not mean that they are more probable to repeat, given by the flat trend.  

```{r user count, message=FALSE, warning=FALSE}
target_vs_colcount(train, "msno", "User Occurence", "Target", "User Occurence vs Target")
```


##### Target is balanced

```{r}
train %>% 
  group_by(target) %>% 
  count
```


#### MEMBERS

```{r read members, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}
members <- as.tibble(fread('/Volumes/EricSeagate/HKUST/Term 2/Machine Learning/Individual Project/members.csv'))
```


```{r display members, echo=FALSE}
members
```

Useful functions,

```{r members function}
members_colgroup <- function(df,col_name, x, y, title, xmin, xmax, ymin, ymax)
{
  
  temp_df <- df %>% 
    group_by_(col_name) %>% 
    count() %>% 
    arrange(desc(n))
  
  df_plot <- temp_df %>% 
    ggplot(aes_string(col_name, "n")) + 
    geom_col(fill='goldenrod2') + 
    labs(x = x,
         y = y,
         title = title) +
    xlim(xmin, xmax) +
    ylim(ymin, ymax) +
    readable_labs
  
  print(df_plot)
  return(temp_df)
  
}

members_date_count <- function(df, col_name, x, y, title)
{
  df %>% 
    group_by_(month = month(col_name), year = year(col_name)) %>% 
    count() %>% 
    ungroup %>% 
    mutate(date = as.Date(paste(as.character(year), as.character(month), '01', sep='-')))
  ggplot(aes(date, n))+
    geom_line(color='goldenrod2', size=1) +
    labs(x = x,
         y = y,
         title= title) +
    xlim(xmin, xmax) +
    readable_labs
}


```

#### Distribution of city, bd(age), gender {.tabset}

##### Age
  


```{r age tab, warning=TRUE}
members_colgroup(members, "bd", "Age", "Frequency", "Age Distribution", 1, 100, 0, 1000)
```

##### City



```{r city tab, warning=TRUE}
members_colgroup(members, "city", "City", "Frequency", "City Distribution", 0, 25, 0, 20000)
```


##### Gender


```{r gender tab, warning=TRUE}
members %>% 
  group_by(gender) %>% 
  count
```


##### Registered_via



```{r reg tab, warning=TRUE}
members_colgroup(members, "registered_via", "Registration Method", "Frequency", "Registration method Distribution", 0, 16, 0, 15000)
```


```{r date conversion}
members %<>% 
  mutate(registration_init_time = ymd(registration_init_time),
         expiration_date = ymd(expiration_date))
```

##### Signup vs Expiration 



```{r signup vs expiration, message=FALSE, warning=FALSE}
#members_date_count(members, "registration_init_time", "Signup Date", "Number of Users", "Signup vs User Count")
reg_count <- members %>% 
  group_by(month = month(registration_init_time), year = year(registration_init_time)) %>% 
  count() %>% 
  ungroup %>% 
  mutate(date = as.Date(paste(as.character(year), as.character(month), '01', sep='-'))) %>% 
  arrange(desc(n)) %>% 
  print

exp_count <- members %>% 
  group_by(month = month(expiration_date), year = year(expiration_date)) %>% 
  count() %>% 
  ungroup %>% 
  mutate(date = as.Date(paste(as.character(year), as.character(month), '01', sep='-'))) %>% 
  arrange(desc(n)) %>% 
  print

reg_count %>% 
  left_join(exp_count, by="date") %>% 
  ggplot() +
  geom_line(aes(date, n.x), color='goldenrod2') +
  geom_line(aes(date, n.y), color='mediumorchid') +
  labs(y="Frequency", title="Registration and Expiration Distribution")+
  readable_labs

```

##### Missingness in members


```{r members missingness}
members %>% 
  mutate(cga = if_else(((city == 1) & (bd == 0) & (gender == "")), 1, 0),
         cg =  if_else(((city == 1) & (gender == "")), 1, 0),
         ca = if_else(((city == 1) & (bd == 0)), 1, 0),
         ga =  if_else(((bd == 0) & (gender == "")), 1, 0)) %>% 
  summarize(city_gender_age = sum(cga),
            city_gender = sum(cg),
            city_age = sum(ca),
            gender_age =sum(ga))
```




#### SONGS

```{r songs read, echo=FALSE, message=FALSE, results=FALSE, warning=FALSE, echo=FALSE}
songs <- as.tibble(fread('/Volumes/EricSeagate/HKUST/Term 2/Machine Learning/Individual Project/songs.csv')) 
```

Songs DF,

```{r songs display}
songs
```

##### Top Items{.tabset}



```{r top 100 function}
top_100 <- function(df, col_name)
{
  temp_df <- df %>% 
    group_by_(col_name) %>% 
    count %>% 
    arrange(desc(n)) %>% 
    print
  
  return(temp_df)
}
```


##### Top 100 Artists

```{r top 100 artist}
artist_count <- top_100(songs, "artist_name")
```

##### Top 100 Lyricist

```{r top 100 lyricist}
lyricist_count <- top_100(songs, "lyricist")
```

##### Top 100 composer

```{r top 100 composer}
composer_count <- top_100(songs, "composer")
```

##### Top 100 Language

```{r top 100 language}
language_count <- top_100(songs, "language")
```


`r sum(songs$artist_name == songs$lyricist)` songs have same artist and lyricist name.  
`r sum(songs$lyricist == songs$composer)` songs have same lyricist and composer name.     
`r sum(songs$artist_name == songs$composer)` songs have same artist and composer name.   
`r sum(songs$artist_name == songs$lyricist)`songs have same artist and lyricist name.  



##### Top Genre's


```{r top genres, message=FALSE, warning=FALSE}
genre_count <- songs %>% 
  separate(genre_ids, c("one", "two", "three", "four", "five", "six", "seven", "eight"), extra="merge") %>% 
  select(one:eight)%>% 
  gather(one:eight, key="nth_id", value="genre_ids", na.rm=TRUE) %>% 
  group_by(genre_ids) %>% 
  count %>% 
  arrange(desc(n)) %>% 
  print()
```

##### Distribution of song length
 

```{r song length dist, message=FALSE, warning=FALSE}
songs %>% 
  mutate(song_length = song_length/6e4) %>% 
  ggplot(aes(song_length)) +
  geom_histogram(binwidth = 0.25, fill='darkorchid3') +
  labs(x='Song Length', y = 'Frequency', title = 'Distribution of song length') +
  xlim(0, 15)
```



#### TEST

```{r read test, message=FALSE, warning=FALSE}
test <- as.tibble(fread('/Volumes/EricSeagate/HKUST/Term 2/Machine Learning/Individual Project/test.csv'))
```

Let's compare the test and train data frames.

##### Distribution of test and train on columns {.tabset}

```{r}
test_train_plot <- function(train, test, col_name, x, y)
  {
  test %>% 
  group_by_(col_name) %>% 
  summarize(count = n()) %>% 
  left_join(train %>% 
              group_by_(col_name) %>% 
              summarize(count = n()) , by=col_name) %>% 
  mutate(ratio = count.x/count.y) %>% 
  rename(test_cnt = count.x, train_cnt = count.y) %>% 
  arrange(ratio) %>% 
  print %>% 
  ggplot() +
  geom_col(aes_string(col_name, "train_cnt"), fill='red', alpha = 0.5) +
  geom_col(aes_string(col_name, "test_cnt"), fill='blue', alpha = 0.5) +
  coord_flip() +
  labs(x = x, y= y)+
  readable_labs
}
```

##### Source system tab

Training set had more records from my library compared to test.

```{r, message=FALSE, warning=FALSE}
test_train_plot(train, test, col_name = "source_system_tab", 'Source system tab', 'Test/Train record Count')
```


##### Source screen name

```{r}
test_train_plot(train, test, col_name = "source_screen_name", "Source Screen Name", "Test/Train Count")
```

##### Source system tab

```{r}
test_train_plot(train, test, col_name = "source_type", "Source Type", "Test/Train Count")
```


### part 2. XGBoost
```{r}
library(reshape)
library(reshape2)
library(xgboost)
library(caret)
library(jsonlite)
library(dplyr)
library(Matrix)
library(doParallel)
library(lubridate)
```


```{r}
train<-read.csv("/Volumes/EricSeagate/HKUST/Term 2/Machine Learning/Individual Project/train.csv")
test<-read.csv("/Volumes/EricSeagate/HKUST/Term 2/Machine Learning/Individual Project/test.csv")

train$key<-paste(train$msno,train$song_id,sep="_")
test$key<-paste(test$msno,test$song_id,sep="_")

train$id<-row.names(train)

test$target<-''

train$type<-'train'
test$type<-'test'

train1<-train[,c('key','type','id','source_system_tab','source_screen_name','source_type','target')]
test1<-test[,c('key','type','id','source_system_tab','source_screen_name','source_type','target')]

master_df<-rbind(train1,test1)

rm(train)
rm(test)
rm(train1)
rm(test1)


#Creating source system tab based primary variables

master_df$flag_source_system_tab_discover<-ifelse(master_df$source_system_tab=="discover",1,0)
master_df$flag_source_system_tab_explore<-ifelse(master_df$source_system_tab=="explore",1,0)
master_df$flag_source_system_tab_listen_with<-ifelse(master_df$source_system_tab=="listen with",1,0)
master_df$flag_source_system_tab_my_library<-ifelse(master_df$source_system_tab=="my library",1,0)
master_df$flag_source_system_tab_notification<-ifelse(master_df$source_system_tab=="notification",1,0)
master_df$flag_source_system_tab_radio<-ifelse(master_df$source_system_tab=="radio",1,0)
master_df$flag_source_system_tab_search<-ifelse(master_df$source_system_tab=="search",1,0)
master_df$flag_source_system_tab_settings<-ifelse(master_df$source_system_tab=="settings",1,0)


print("primary variables")
#Creating source type based primary variables

master_df$flag_source_type_song<-ifelse(master_df$source_type=="song",1,0)
master_df$flag_source_type_song_based_playlist<-ifelse(master_df$source_type=="song-based-playlist",1,0)
master_df$flag_source_type_top_hits_for_artist<-ifelse(master_df$source_type=="top-hits-for-artist",1,0)
master_df$flag_source_type_topic_article_playlist<-ifelse(master_df$source_type=="topic-article-playlist",1,0)
master_df$flag_source_type_my_daily_playlist<-ifelse(master_df$source_type=="my-daily-playlist",1,0)
master_df$flag_source_type_online_playlist<-ifelse(master_df$source_type=="online-playlist",1,0)
master_df$flag_source_type_listen_with<-ifelse(master_df$source_type=="listen-with",1,0)
master_df$flag_source_type_local_library<-ifelse(master_df$source_type=="local-library",1,0)
master_df$flag_source_type_local_playlist<-ifelse(master_df$source_type=="local-playlist",1,0)
master_df$flag_source_type_album<-ifelse(master_df$source_type=="album",1,0)
master_df$flag_source_type_artist<-ifelse(master_df$source_type=="artist",1,0)



#Creating secondary variables combining source type and source tab information
#For album
master_df$flag_source_type_system_tab_album_discover<-master_df$flag_source_type_album*master_df$flag_source_system_tab_discover
master_df$flag_source_type_system_tab_album_my_library<-master_df$flag_source_type_album*master_df$flag_source_system_tab_my_library
master_df$flag_source_type_system_tab_album_search<-master_df$flag_source_type_album*master_df$flag_source_system_tab_search

#For artist
master_df$flag_source_type_system_tab_artist_my_library<-master_df$flag_source_type_artist*master_df$flag_source_system_tab_my_library

#For listen with
master_df$flag_source_type_system_tab_listen_with<-master_df$flag_source_system_tab_listen_with*master_df$flag_source_type_listen_with

#For local library
master_df$flag_source_type_system_tab_local_library_my_library<-master_df$flag_source_type_local_library*master_df$flag_source_system_tab_my_library
master_df$flag_source_type_system_tab_local_library_discover<-master_df$flag_source_type_local_library*master_df$flag_source_system_tab_discover

#For local playlist
master_df$flag_source_type_system_tab_local_playlist_my_library<-master_df$flag_source_type_local_playlist*master_df$flag_source_system_tab_my_library

#For online playlist
master_df$flag_source_type_system_tab_online_playlist_discover<-master_df$flag_source_type_online_playlist*master_df$flag_source_system_tab_discover

print("secondary variables")
#For song
master_df$flag_source_type_system_tab_song_search<-master_df$flag_source_type_song*master_df$flag_source_system_tab_search

#For song based playlist
master_df$flag_source_type_system_tab_song_based_playlist_discover<-master_df$flag_source_type_song_based_playlist*master_df$flag_source_system_tab_discover

train_data<-master_df[master_df$type=='train',]
test_data<-master_df[master_df$type=='test',]

#training and test data creation
train_data_xgb1<-master_df[master_df$type=="train",c(8:ncol(master_df))]
test_data_xgb1<-master_df[master_df$type=="test",c(8:ncol(master_df))]

train_data$target <- factor(train_data$target, levels = c(0,1), ordered = TRUE)
ydata <- as.numeric(train_data$target)-1

xdata <- Matrix(as.matrix(train_data_xgb1), sparse = TRUE)

xdata_test_final <- Matrix(as.matrix(test_data_xgb1), sparse = TRUE)

#####################Analysis done########################

#####################XGBoost start########################
# xgboost parameters
param <- list("objective" = "multi:softprob",    # multiclass classification 
              "num_class" = 2,    # number of classes 
              "eval_metric" = "merror",    # evaluation metric 
              "nthread" = 8,   # number of threads to be used 
              "max_depth" = 16,    # maximum depth of tree 
              "eta" = 0.3,    # step size shrinkage 
              "gamma" = 0,    # minimum loss reduction 
              "subsample" = 1,    # part of data instances to grow tree 
              "colsample_bytree" = 1,  # subsample ratio of columns when constructing each tree 
              "min_child_weight" = 12  # minimum sum of instance weight needed in a child 
              )

print("XGBoost")
bst.cv <- xgb.cv(param=param, data=xdata, label=ydata, 
              nfold=2, nrounds=30, prediction=TRUE, verbose=TRUE)

min.merror.idx = which.min(bst.cv$evaluation_log[,test_merror_mean]) 

xgb <- xgboost(param=param, data=xdata, label=ydata,
                           nrounds=min.merror.idx, verbose=TRUE)


pred_xgb <- predict(xgb, xdata_test_final, reshape = TRUE)


pred_xgb2 <- as.data.frame(pred_xgb)
names(pred_xgb2) <- c("zero","one")
pred_xgb2$id <- test_data$id

pred_xgb2$target <- ifelse(pred_xgb2$zero>pred_xgb2$one,0,1)

pred_xgb3<-pred_xgb2[,c(3,4)]
print("Export prediction")
write.csv(pred_xgb3, "predictions.csv", row.names = FALSE)
```
